---
title: "Proposal data"
author: "Ellie and Finn"
output: github_document
---

```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
```

```{r}
data.rough <- read.csv("veg_data_rough.csv")
```

## veg_data_rough

```{r}
head(data.rough
     )
```

- `location`: qualitative character variable; the name of the island the data observation was collected on
- `collected.by`: qualitative character variable; who was taking the observations
- `data`: discrete numerical; date data was recorded
- `gps`: qualitative character variable; plot code as denoted in ArcGIS/GPS unit
- `cover`: discrete ordinal; % cover classes (entered as highest number in range)
- `height`: discrete ordinal; height classes up to >1.5m, entered as their class code (height range information in metadata, number denotes which class. Height increases as code increases.)
- `common`: qualitative character variable; common name for species observed.
- `spp_code`: qualitative character variable; 6-letter species code
- `spp`: qualitative character variable; scientific name for species observed
- `notes`: qualitative character variable; notes from plot

Each row is a species observation.
- ...
