---
title: "Data analysis"
author: "Ellie and Finn"
output:
  html_document:
    df_print: paged
---
```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(leaflet) ## For leaflet interactive maps
library(sf) ## For spatial data
library(RColorBrewer) ## For colour palettes
library(htmltools) ## For html
library(leafsync) ## For placing plots side by side
library(stringr) ## For character string manipulation
library(kableExtra) ## Table  output (in slides)
#install.packages("visdat")
library(visdat)
# install.packages("naniar")
library(naniar)
library(readxl) ##working with xcel sheets
library(ggridges)
```

### READ IN DATA
```{r read-in-vegetation-file, echo = FALSE}
island_veg <- read_csv("../data/island_veg.csv")
veg_attributes <- read_csv("../data/plotpointsattributes.csv") 
gdi_attributes <- read_csv("../data/gdipointsattributes.csv")
```
-
```{r fix-gdi-attributes}
gdi_attributes <- gdi_attributes %>%
  mutate(name = paste0('gdi',OID_))
```
-
```{r join-gdi}
veg_attributes <- full_join(veg_attributes, gdi_attributes)
```
-
```{r read-in-bird-files, echo = FALSE}
gdi_bird <- st_read("../data/gull data/gdi_nests/gdi_nests_Merge.shp")
heron_bird <- st_read("../data/gull data/heron_nests/herg_nests_heron.shp")
#coei_schoodic_bird <- st_read("../data/gull data/schoodic_nests/coei_nests.shp")
#gbbg_schoodic_bird <- st_read("../data/gull data/schoodic_nests/gbbg_nests.shp")
#herg_schoodic_bird <- st_read("../data/gull data/schoodic_nests/herg_nests.shp")
schoodic_bird <- st_read("../data/gull data/schoodic_nests/all_nests.shp")
shabby_bird <- st_read("../data/gull data/shabby_nests/gull_eider_nests.shp")
```

-
###TIDYING

##Tidy bird data
```{r create-location-colmn-birds}
gdi_bird$location <- c("gdi")
heron_bird$location <- c("heron")
shabby_bird$location <- c("shabby")
schoodic_bird$location <- c("schoodic")
```
-
```{r recode-shabby-column-names}
  colnames(shabby_bird)[which(names(shabby_bird) == "Point_ge_4")] <- "GPS_Date"
  colnames(shabby_bird)[which(names(shabby_bird) == "shabby_i_2")] <- "species"
  colnames(shabby_bird)[which(names(shabby_bird) == "Point_gene")] <- "Point_ID"
```
-
```{r recode-schoodic-column-names}
  colnames(schoodic_bird)[which(names(schoodic_bird) == "all_nest_3")] <- "GPS_Date"
  colnames(schoodic_bird)[which(names(schoodic_bird) == "schoodic_2")] <- "species"
  colnames(schoodic_bird)[which(names(schoodic_bird) == "schoodic_c")] <- "Point_ID"
```
-
```{r select-relevant-columns-in-bird-data}
gdi_bird_all <- gdi_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)
 
heron_bird_all <- heron_bird %>%
  mutate(species = "herring gull") %>% #create species column for heron
  select(Point_ID, GPS_Date, species, location, geometry)
  
shabby_bird_all <- shabby_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)

schoodic_bird_all <- schoodic_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)
```
-
```{r combine-bird-datasets-for-all-islands}
island_bird <- rbind(gdi_bird_all, heron_bird_all, shabby_bird_all, schoodic_bird_all)
```
-
```{r rename-bird-species-with-full-names}
island_bird <- island_bird %>%
  mutate(species = case_when(
    species == "gbbg" ~ "great black-backed gull",
    species == "herg" ~ "herring gull",
    species == "coei" ~ "common eider",
    species == "dcco" ~ "double-crested cormorant",
    species == "cago" ~ "canada goose",
    TRUE ~ species
  )) %>% 
  filter(!is.na(species)) %>% 
  filter(species != "NA")
```
-
```{r visualize-missing-bird-data}
island_bird_dataframe <- as.data.frame(island_bird)
visdat::vis_dat(island_bird_dataframe)
```
#create bird color palette
```{r assign-colors-to-bird-species}
 colors <- c("herring gull" = "#21918c", 
             "great black-backed gull" = "#440154", 
             "common eider" = "#fde725", 
             "canada goose" = "#5ec962", 
             "double-crested cormorant" = "#3b528b")
```
-
```{r assign-colors-to-bird-species-for-leaflet}
 bird_color <- colorFactor(palette = c("#21918c", "#440154", "#fde725", "#5ec962", "#3b528b"),
                      domain = c("herring gull", "great black-backed gull", "common eider", "canada goose", "double-crested cormorant"), 
                      ordered = TRUE)
```
-

##Tidy vegetation data
```{r join veg & coordinates}
island_veg_all <- island_veg %>%
                  left_join(veg_attributes, join_by("gps" == "name"))
```
-
```{r standardize-island-naming}
island_veg_all <- island_veg_all %>%
  mutate(location = case_when(
    location == "Schoodic" ~ "schoodic",
    location == "GDI" ~ "gdi",
    TRUE ~ location
  ))
```
-
```{r recode-veg-height-class-names}
 island_veg_all <- island_veg_all %>%
mutate(height_class = case_when(TRUE ~ as.factor(height))) %>% #making height class a factor variable in order to use it in a bar chart
  mutate(height_class_range = case_when(
    height_class == "1" ~ "0-0.15",
    height_class == "2" ~ "0.15-0.30",
    height_class == "3" ~ "0.30-1",
    height_class == "4" ~ "1-1.5",
    height_class == "5" ~ "1.5+"
  ))
```
-
```{r recode-veg-cover-names}
island_veg_all <- island_veg_all %>%
  mutate(percent_cover_range = case_when(
    cover == "1" ~ "0-0.01",
    cover == "5" ~ "0.02-0.05",
    cover == "25" ~ "0.06-0.25",
    cover == "50" ~ "0.26-0.50",
    cover == "75" ~ "0.51-0.75",
    cover == "100" ~ "0.76-1.00"
  ))
```
-
```{r standardize spp (veg)}
island_veg_all <- island_veg_all %>% 
  mutate(spp = case_when(
    spp == "achillea millefolium" ~ "Achillea millefolium",
    spp == "agrostis scabra" ~ "Agrostis scabra",
    spp == "agrostis stolonifera" ~ "Agrostis stolonifera",
    spp == "arctium sp" ~ "Arctium sp.",
    spp == "calystegia sepium" ~ "Calystegia sepium",
    spp == "carex nigra" ~ "Carex nigra",
    spp == "cerastium arvense" ~ "Cerastium arvense",
    spp == "cerastium fontana" ~ "Cerastium fontanum",
    spp == "cerastium fontanum" ~ "Cerastium fontanum",
    spp == "cirsium arvense" ~ "Cirsium arvense",
    spp == "cirsium vulgare" ~ "Cirsium vulgare",
    spp == "cornus canadensis" ~ "Cornus canadensis",
    spp == "cuscuta gronovii" ~ "Cuscuta gronovii",
    spp == "doellingeria umbellata" ~ "Doellingeria umbellata",
    spp == "festuca rubra" ~ "Festuca rubra",
    spp == "fragaria virginiana" ~ "Fragaria virginiana",
    spp == "hieracium caespitosum" ~ "Hieracium caespitosum",
    spp == "impatiens capensis" ~ "Impatiens capensis",
    spp == "iris versicolor" ~ "Iris versicolor",
    spp == "lathyrus japonicus" ~ "Lathyrus japonicus",
    spp == "lathyrus palustris" ~ "Lathyrus palustris",
    spp == "luzula multiflora" ~ "Luzula multiflora",
    spp == "maianthemum canadense" ~ "Maianthemum canadense",
    spp == "matricaria discoidea" ~ "Matricaria discoidea",
    spp == "myrica gale" ~ "Myrica gale",
    spp == "osmundastrum cinnamomeum" ~ "Osmundastrum cinnamomeum",
    spp == "plantago major" ~ "Plantago major",
    spp == "poa pratensis" ~ "Poa pratensis",
    spp == "Poa spp (tufting)" ~ "Poa sp.",
    spp == "potentilla anserina" ~ "Potentilla anserina",
    spp == "potentilla norvegica" ~ "Potentilla norvegica",
    spp == "potentilla simplex" ~ "Potentilla simplex",
    spp == "ranunculus acris" ~ "Ranunculus acris",
    spp == "rubus hispidus" ~ "Rubus hispidus",
    spp == "rubus idaeus" ~ "Rubus idaeus",
    spp == "rumex acetosella" ~ "Rumex acetosella",
    spp == "rumex longifolius" ~ "Rumex longifolius",
    spp == "rumex sp." ~ "Rumex sp.",
    spp == "scutellaria galericulata" ~ "Scutellaria galericulata",
    spp == "solanum dulcamara" ~ "Solanum dulcamara",
    spp == "solidago rugosa" ~ "Solidago rugosa",
    spp == "solidago sempervirens" ~ "Solidago sempervirens",
    spp == "spiraea alba" ~ "Spiraea alba",
    spp == "stellaria media" ~ "Stellaria media",
    spp == "Strawberry" ~ "Fragaria virginiana",
    spp == "symphotrichum novae-angliae" ~ "Symphotrichum novae-angliae",
    spp == "taraxacum officinale" ~ "Taraxacum officinale",
    spp == "taraxacum officionale" ~ "Taraxacum officinale",
    spp == "trifolium repens" ~ "Trifolium repens",
    spp == "urtica dioica" ~ "Urtica dioica",
    spp == "vaccinium angustifolium" ~ "Vaccinium angustifolium",
    spp == "valeriana officionalis" ~ "Valeriana officinalis",
    spp == "Valeriana officionalis" ~ "Valeriana officinalis",
    spp == "valeriana officinalis" ~ "Valeriana officinalis",
    spp == "veronica arvensis" ~ "Veronica arvensis",
    TRUE ~ spp
  ))
print(island_veg_all)
```
-
```{r insert IDs (veg)}
island_veg_all <- island_veg_all %>% 
  mutate(spp = case_when(
    spp == "Min lat" ~ "Moehringia lateriflora",
    spp == "Calamagrostis #8" ~ "Poa pratensis",
    spp == "grass #54" ~ "Poa pratensis",
    spp == "Poa pratensis (#16)" ~ "Poa pratensis",
    spp == "Poa pratensis #16)" ~ "Poa pratensis",
    spp == "Poa spp #19" ~ "Alopecurus pratensis",
    spp == "elymus #33" ~ "Elymus trachycaulus",
    spp == "grass #74" ~ "Agrostis stolonifera",
    spp == "Poa #3" ~ "Poa trivialis",
    spp == "#53" ~ "Capsella bursa-pastoris",
    spp == "Poaceae #84" ~ "Elymus repens",
    spp == "Sedge #82" ~ "Carex silicea",
    spp == "Dock sp #45" ~ "Rumex britannica",
    spp == "dock sp #51" ~ "Rumex crispa",
    spp == "Poaceae #22" ~ "Anthoxanthum odoratum",
    spp == "Vetch #99" ~ "Lathyrus palustris",
    TRUE ~ spp
  ))
```
-
```{r create-veg-types}
island_veg_all$type <- "NA"
```
-
```{r assign-veg-types}
island_veg_all <- island_veg_all %>%
  # select(location:percent_cover_range) %>%
   mutate(type = case_when(
    spp == "Achillea millefolium" ~ "herb",
    spp == "Agrostis scabra" ~ "gram",
    spp == "Agrostis stolonifera" ~ "gram",
    spp == "Arctium sp." ~ "herb",
    spp == "Calystegia sepium" ~ "vine",
    spp == "Carex nigra" ~ "gram",
    spp == "Cerastium arvense" ~ "herb",
    spp == "Cerastium fontanum" ~ "herb",
    spp == "Cerastium fontanum" ~ "herb",
    spp == "Cirsium arvense" ~ "herb",
    spp == "Cirsium vulgare" ~ "herb",
    spp == "Cornus canadensis" ~ "herb",
    spp == "Cuscuta gronovii" ~ "vine",
    spp == "Doellingeria umbellata" ~ "herb",
    spp == "Festuca rubra" ~ "gram",
    spp == "Fragaria virginiana" ~ "herb",
    spp == "Hieracium caespitosum" ~ "herb",
    spp == "Impatiens capensis" ~ "herb",
    spp == "Iris versicolor" ~ "herb",
    spp == "Lathyrus japonicus" ~ "vine",
    spp == "Lathyrus palustris" ~ "vine",
    spp == "Luzula multiflora" ~ "gram",
    spp == "Maianthemum canadense" ~ "herb",
    spp == "Matricaria discoidea" ~ "herb",
    spp == "Myrica gale" ~ "shrub",
    spp == "Osmundastrum cinnamomeum" ~ "herb",
    spp == "Plantago major" ~ "herb",
    spp == "Poa pratensis" ~ "gram",
    spp == "Poa sp." ~ "gram",
    spp == "Potentilla anserina" ~ "herb",
    spp == "Potentilla norvegica" ~ "herb",
    spp == "Potentilla simplex" ~ "herb",
    spp == "Ranunculus acris" ~ "herb",
    spp == "Rubus hispidus" ~ "vine",
    spp == "Rubus idaeus" ~ "shrub",
    spp == "Rumex acetosella" ~ "herb",
    spp == "Rumex longifolius" ~ "herb",
    spp == "Rumex sp." ~ "herb",
    spp == "Rubus shrub spp. (#32)" ~ "shrub",
    spp == "Scutellaria galericulata" ~ "herb",
    spp == "Solanum dulcamara" ~ "vine",
    spp == "Solidago rugosa" ~ "herb",
    spp == "Solidago sempervirens" ~ "herb",
    spp == "Spiraea alba" ~ "shrub",
    spp == "Stellaria media" ~ "herb",
    spp == "Symphotrichum novae-angliae" ~ "herb",
    spp == "Taraxacum officinale" ~ "herb",
    spp == "Trifolium repens" ~ "herb",
    spp == "Urtica dioica" ~ "herb",
    spp == "Vaccinium angustifolium" ~ "shrub",
    spp == "Valeriana officinalis" ~ "herb",
    spp == "Veronica arvensis" ~ "herb",
    TRUE ~ type
  ))
```
#visualize the dataset
```{r find-unique-species-veg}
unique <- unique(island_veg_all$spp)
sort(unique)
```
-
```{r visualize-missing-veg-data}
visdat::vis_dat(island_veg_all)
```
-
```{r visualize-percent-missing-veg-data}
visdat::vis_miss(island_veg_all)
```
-
```{r visualize-missing-variables-veg-data}
naniar::gg_miss_var(island_veg_all)
```
#create veg shape file from csv
```{r create-veg-shape-file}
# #removing columns that are entirely na
# island_veg_no_na <- island_veg_all %>%
#   select(location, date, gps, cover, height, common, spp_code, spp, notes, POINT_X, POINT_Y, height_class, height_class_range, percent_cover_range, type)
#   
# #removing gdi, because it has no coordinates
# island_veg_no_na_2 <- filter(island_veg_no_na, location != "gdi")
# 
# #creating a vegetation shape file
# island_veg_shape <- st_as_sf(island_veg_no_na_2, coords = c(10:11))

island_veg_shape <- st_as_sf(island_veg_all, coords = c(13:14))

```
<<<<<<< HEAD
#create veg color palettes
```{r assign-colors-to-percent-cover}
cover_color <- c("0-0.01" = "#dadaeb", 
                 "0.06-0.25" = "#bcbddc", 
                 "0.26-0.50" = "#9e9ac8", 
                 "0.51-0.75" = "#756bb1", 
                 "0.76-1.00" = "#54278f") 
```
-
```{r assign-colors-to-percent-cover-for-leaflet}
cover_color_leaf <- colorFactor(palette = c("#dadaeb", "#bcbddc", "#9e9ac8", "#756bb1", "#54278f"), 
                           domain = c("0-0.01", "0.06-0.25", "0.26-0.50", "0.51-0.75", "0.76-1.00"),
                           ordered = TRUE)
```
-
```{r assign-colors-to-height-classes}
height_color <- c("0-0.15" = "#ffffb2", 
                  "0.15-0.30" = "#fecc5c", 
                  "0.30-1" = "#fd8d3c", 
                  "1-1.5" = "#f03b20", 
                  "1.5+" = "#bd0026")
```
-
```{r assign-colors-to-height-classes-for-leaflet}
height_color_leaf <- colorFactor(palette = c("#ffffb2", "#fecc5c", "#fd8d3c", "#f03b20", "#bd0026"), 
                            domain = c("0-0.15", "0.15-0.30", "0.30-1", "1-1.5", "1.5+"),
                            ordered = TRUE)
```
-
=======
>>>>>>> 9943bb3465ffeb17203d46eada48e91b9adb692e

###VISUALIZING

##Visualize bird data
```{r bird-species-bar-plot-by-island}
#I'm trying to figure out how to display totals on the ends of the bars so you can see numbers that don't show up in the bars. 
island_bird %>%
  ggplot(mapping = aes(y = species),
  na.rm = TRUE) +
  geom_bar(fill = "#21918c") +
  facet_wrap(~location, nrow = 1) +
  theme(axis.text.y = element_text(vjust = 10)) +
  theme_minimal() + 
  labs(title = "Seabird disribution by island",
       x = "Number of occurences",
       y = "Bird species")
```
-
```{r species-percentage-by-island}
island_bird %>%
  ggplot(mapping = aes(x = location, fill = (species))) + 
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Bird species on each island",
       x = "Island",
       y = "Proportion of species",
       fill = "Bird species") +
  theme_minimal()
```
#geom_sf maps of species on each island
```{r ggplot-map-of-nests-on-gdi, fig.width=20, dev='jpeg'}
island_bird %>%
filter(location == "gdi") %>% #filter for one island for better vis
ggplot() +
  geom_sf(aes(color = species)) + #spacial geom
  scale_color_manual(values = colors) + #use same color palette so species match on each plot
  labs(title = "Seabird nests on Great Duck Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-heron, fig.width=20, dev='jpeg'}
island_bird %>%
filter(location == "heron") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-schoodic, fig.width=20, dev='jpeg'}
island_bird %>%
filter(location == "schoodic") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-shabby, fig.width=20, dev='jpeg'}
island_bird %>%
filter(location == "shabby") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-

##Visualize vegetation data
```{r species-barplot-by-island, fig.height=10, dev='jpeg'}
island_veg_all %>%
  filter(spp != "unknown") %>% #filtering out all of the unknown species
  ggplot(mapping = aes(y = spp)
    ) +
  geom_bar(fill = "#21918c") +
  facet_wrap(~location, nrow = 1) +
  theme(axis.text.y = element_text(vjust = 10)) +
  theme_minimal() + 
  labs(title = "Vegetation frequency by island",
       x = "Number of occurences",
       y = "Scientific names")


# png("island_plot.png", width = 480, height = 480)
# island_veg %>%
#   filter(common != "unknown") %>% #filtering out all of the unknown species
#   ggplot(mapping = aes(y = common)
#     ) +
#   geom_bar(fill = "#21918c") +
#   facet_wrap(~location, nrow = 1) +
#   theme(axis.text.y = element_text(vjust = 10)) +
#   theme_minimal() + 
#   labs(title = "Species disribution by island",
#        x = "Number of occurences",
#        y = "Common names")
# dev.off()
```
-
```{r height-class-distribution-by-island-bar-plot}
island_veg_all %>%
  ggplot(mapping = aes(x = location, fill = fct_rev(height_class_range))) + #reversing height class so the tallest appears on top
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Height classes on each island",
       x = "Island",
       y = "Proportion on height class",
       fill = "Plant height classes (m)") +
  theme_minimal()
```
-
```{r not-working-yet-veg-types-by-island}
island_veg_all %>% 
  ggplot(mapping = aes(x = location, fill = type)) +
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Veg types on each island",
       x = "Island",
       y = "Proportion on height class",
       fill = "Plant types") +
  theme_minimal()
```
#geom_point maps of the percent cover on each island
```{r ggplot-geom_point-map-percent-cover-on-heron}
island_veg_all %>%
  filter(location == "heron") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) + #geom point because data doesn't read as geometry. Rename column?
  scale_color_manual(values = cover_color) + #color scale for percent cover, darker is more cover
  scale_y_continuous(labels = ~ paste0(.x, "°N")) + #add lat/long symbols to axis
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
-
```{r ggplot-geom_point-map-percent-cover-on-schoodic}
island_veg_all %>%
  filter(location == "schoodic") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) +
  scale_color_manual(values = cover_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
-
```{r ggplot-geom_point-map-percent-cover-on-shabby}
island_veg_all %>%
  filter(location == "shabby") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) +
  scale_color_manual(values = cover_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
#geom_point maps of the height classes on each island
```{r ggplot-geom_point-map-height-on-heron}
island_veg_all %>%
  filter(location == "heron") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) + #rev height range so it goes low to high
  scale_color_manual(values = height_color) + #special color scale for height
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-
```{r ggplot-geom_point-map-height-on-schoodic}
island_veg_all %>%
  filter(location == "schoodic") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) +
  scale_color_manual(values = height_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-
```{r ggplot-geom_point-map-height-on-shabby}
island_veg_all %>%
  filter(location == "shabby") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) +
  scale_color_manual(values = height_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-

##Compare birds and vegetation
```{r change_coordinate_system_for_bird}
#update bird data to the coordinate system used in leaflet
island_bird <- st_transform(island_bird, "+init=epsg:4326")
```
-
```{r leaflet-map-of-height-and-bird-species}
leaflet() %>% 
  #basemap
  addProviderTiles("CartoDB.Positron") %>% 
  #setView to show all islands
  setView(lng = -68.25,
          lat = 44.25,
          zoom = 10) %>% 
  #add markers for the birds and color by species using bird_color palette
  addCircleMarkers(data = island_bird,
                   color = "black",
                   weight = 1, #~bird_color(island_bird$species),
                   fillColor = ~bird_color(island_bird$species),
                   radius = 3,
                   opacity = .5,
                   fillOpacity = 1) %>% 
  #add markers for veg colored by height class using the height_color_leaf palette
  addCircleMarkers(data = island_veg_shape, 
                   color = ~height_color_leaf(island_veg_shape$height_class_range), 
                   fillColor = ~height_color_leaf(island_veg_shape$height_class_range),
                   radius = 2,
                   opacity = 1,
                   fillOpacity = 1)
    
```

#I would like to have a representation of bird densities and then veg plots colored by height
#also veg plots represented by dominant vegetation type in colors
```{r species richness representations}
#ridge plot of # of species in each plot by island
iva <- island_veg_all %>%
  group_by(gps) %>%
  mutate(spp = replace_na(spp, "1")) %>%
   mutate(numSp = n_distinct(spp)) %>%
   mutate(NSp = if_else(spp != 1, numSp, 0)) %>%
  select(-numSp) %>%
  aggregate(NSp~gps+location, FUN = "mean")

ggplot(iva, aes(NSp, location, fill = fct_rev(location))) +
  geom_density_ridges(alpha = 0.5, bandwidth = .6) +
  xlim(0,15) +
  scale_fill_viridis_d() +
  labs(title = "Species Richness per Plot, by Island",
       y = "Island",
       x = "# of Species/Plot",
       fill = "Island"
  )
```


