---
title: "Data analysis"
author: "Ellie and Finn"
output:
  html_document:
    df_print: paged
---
```{r load-packages, message = FALSE}
library(tidyverse)
library(broom)
library(leaflet) ## For leaflet interactive maps
library(sf) ## For spatial data
library(RColorBrewer) ## For colour palettes
library(htmltools) ## For html
library(leafsync) ## For placing plots side by side
library(stringr) ## For character string manipulation
library(kableExtra) ## Table  output (in slides)
#install.packages("visdat")
library(visdat)
# install.packages("naniar")
library(naniar)
library(readxl) ##working with xcel sheets
library(ggridges)
```

### READ IN DATA
```{r read-in-vegetation-file, echo = FALSE}
island_veg <- read_csv("../data/island_veg.csv")
veg_attributes <- read_csv("../data/plotpointsattributes.csv") 
gdi_attributes <- read_csv("../data/gdipointsattributes.csv")
```
-
```{r fix-gdi-attributes}
gdi_attributes <- gdi_attributes %>%
  mutate(name = paste0('gdi',OID_))
```
-
```{r join-gdi}
veg_attributes <- full_join(veg_attributes, gdi_attributes)
```
-
```{r read-in-bird-files, echo = FALSE}
gdi_bird <- st_read("../data/gull data/gdi_nests/gdi_nests_Merge.shp")
heron_bird <- st_read("../data/gull data/heron_nests/herg_nests_heron.shp")
#coei_schoodic_bird <- st_read("../data/gull data/schoodic_nests/coei_nests.shp")
#gbbg_schoodic_bird <- st_read("../data/gull data/schoodic_nests/gbbg_nests.shp")
#herg_schoodic_bird <- st_read("../data/gull data/schoodic_nests/herg_nests.shp")
schoodic_bird <- st_read("../data/gull data/schoodic_nests/all_nests.shp")
shabby_bird <- st_read("../data/gull data/shabby_nests/gull_eider_nests.shp")
```

-
###TIDYING

#Tidy bird data
```{r create-location-colmn-birds}
gdi_bird$location <- c("gdi")
heron_bird$location <- c("heron")
shabby_bird$location <- c("shabby")
schoodic_bird$location <- c("schoodic")
```
-
```{r recode-shabby-column-names}
  colnames(shabby_bird)[which(names(shabby_bird) == "Point_ge_4")] <- "GPS_Date"
  colnames(shabby_bird)[which(names(shabby_bird) == "shabby_i_2")] <- "species"
  colnames(shabby_bird)[which(names(shabby_bird) == "Point_gene")] <- "Point_ID"
```
-
```{r recode-schoodic-column-names}
  colnames(schoodic_bird)[which(names(schoodic_bird) == "all_nest_3")] <- "GPS_Date"
  colnames(schoodic_bird)[which(names(schoodic_bird) == "schoodic_2")] <- "species"
  colnames(schoodic_bird)[which(names(schoodic_bird) == "schoodic_c")] <- "Point_ID"
```
-
```{r select-relevant-columns-in-bird-data}
gdi_bird_all <- gdi_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)
 
heron_bird_all <- heron_bird %>%
  mutate(species = "herring gull") %>% #create species column for heron
  select(Point_ID, GPS_Date, species, location, geometry)
  
shabby_bird_all <- shabby_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)

schoodic_bird_all <- schoodic_bird %>%
  select(Point_ID, GPS_Date, species, location, geometry)
```
-
```{r combine-bird-datasets-for-all-islands}
island_bird <- rbind(gdi_bird_all, heron_bird_all, shabby_bird_all, schoodic_bird_all)
```
-
```{r rename-bird-species-with-full-names}
island_bird <- island_bird %>%
  mutate(species = case_when(
    species == "gbbg" ~ "great black-backed gull",
    species == "herg" ~ "herring gull",
    species == "coei" ~ "common eider",
    species == "dcco" ~ "double-crested cormorant",
    species == "cago" ~ "canada goose",
    TRUE ~ species
  )) %>% 
  filter(!is.na(species)) %>% 
  filter(species != "NA")
```
-
```{r visualize-missing-bird-data, fig.alt = "Colored block showing missing values in the island_bird data where no values are missing"}
island_bird_dataframe <- as.data.frame(island_bird)
visdat::vis_dat(island_bird_dataframe)
```
##create bird color palette
```{r assign-colors-to-bird-species}
#Using the Paul Tol pallete to be colorblind friendly
 colors <- c("herring gull" = "#88CCEE", 
             "great black-backed gull" = "#332288", 
             "common eider" = "#CC6677", 
             "canada goose" = "#117733", 
             "double-crested cormorant" = "#882255")
```
-
```{r assign-colors-to-bird-species-for-leaflet}
#Using the Paul Tol pallete to be colorblind friendly
 bird_color <- colorFactor(palette = c("#88CCEE", "#332288", "#CC6677", "#117733", "#882255"),
                      domain = c("herring gull", "great black-backed gull", "common eider", "canada goose", "double-crested cormorant"), 
                      ordered = TRUE)
```
-

#Tidy vegetation data
```{r join veg & coordinates}
island_veg_all <- island_veg %>%
                  left_join(veg_attributes, join_by("gps" == "name"))
```
-
```{r standardize-island-naming}
island_veg_all <- island_veg_all %>%
  mutate(location = case_when(
    location == "Schoodic" ~ "schoodic",
    location == "GDI" ~ "gdi",
    TRUE ~ location
  ))
```
-
```{r recode-veg-height-class-names}
 island_veg_all <- island_veg_all %>%
mutate(height_class = case_when(TRUE ~ as.factor(height))) %>% #making height class a factor variable in order to use it in a bar chart
  mutate(height_class_range = case_when(
    height_class == "1" ~ "0-0.15",
    height_class == "2" ~ "0.15-0.30",
    height_class == "3" ~ "0.30-1",
    height_class == "4" ~ "1-1.5",
    height_class == "5" ~ "1.5+"
  ))
```
-
```{r recode-veg-cover-names}
island_veg_all <- island_veg_all %>%
  mutate(percent_cover_range = case_when(
    cover == "1" ~ "0-0.01",
    cover == "5" ~ "0.02-0.05",
    cover == "25" ~ "0.06-0.25",
    cover == "50" ~ "0.26-0.50",
    cover == "75" ~ "0.51-0.75",
    cover == "100" ~ "0.76-1.00"
  ))
```
-
```{r standardize spp (veg)}
island_veg_all <- island_veg_all %>% 
  mutate(spp = case_when(
    spp == "achillea millefolium" ~ "Achillea millefolium",
    spp == "agrostis scabra" ~ "Agrostis scabra",
    spp == "agrostis stolonifera" ~ "Agrostis stolonifera",
    spp == "arctium sp" ~ "Arctium sp.",
    spp == "calystegia sepium" ~ "Calystegia sepium",
    spp == "carex nigra" ~ "Carex nigra",
    spp == "cerastium arvense" ~ "Cerastium arvense",
    spp == "cerastium fontana" ~ "Cerastium fontanum",
    spp == "cerastium fontanum" ~ "Cerastium fontanum",
    spp == "cirsium arvense" ~ "Cirsium arvense",
    spp == "cirsium vulgare" ~ "Cirsium vulgare",
    spp == "cornus canadensis" ~ "Cornus canadensis",
    spp == "cuscuta gronovii" ~ "Cuscuta gronovii",
    spp == "doellingeria umbellata" ~ "Doellingeria umbellata",
    spp == "festuca rubra" ~ "Festuca rubra",
    spp == "fragaria virginiana" ~ "Fragaria virginiana",
    spp == "hieracium caespitosum" ~ "Hieracium caespitosum",
    spp == "impatiens capensis" ~ "Impatiens capensis",
    spp == "iris versicolor" ~ "Iris versicolor",
    spp == "lathyrus japonicus" ~ "Lathyrus japonicus",
    spp == "lathyrus palustris" ~ "Lathyrus palustris",
    spp == "luzula multiflora" ~ "Luzula multiflora",
    spp == "maianthemum canadense" ~ "Maianthemum canadense",
    spp == "matricaria discoidea" ~ "Matricaria discoidea",
    spp == "myrica gale" ~ "Myrica gale",
    spp == "osmundastrum cinnamomeum" ~ "Osmundastrum cinnamomeum",
    spp == "plantago major" ~ "Plantago major",
    spp == "poa pratensis" ~ "Poa pratensis",
    spp == "Poa spp (tufting)" ~ "Poa sp.",
    spp == "potentilla anserina" ~ "Potentilla anserina",
    spp == "potentilla norvegica" ~ "Potentilla norvegica",
    spp == "potentilla simplex" ~ "Potentilla simplex",
    spp == "ranunculus acris" ~ "Ranunculus acris",
    spp == "rubus hispidus" ~ "Rubus hispidus",
    spp == "rubus idaeus" ~ "Rubus idaeus",
    spp == "rumex acetosella" ~ "Rumex acetosella",
    spp == "rumex longifolius" ~ "Rumex longifolius",
    spp == "rumex sp." ~ "Rumex sp.",
    spp == "scutellaria galericulata" ~ "Scutellaria galericulata",
    spp == "solanum dulcamara" ~ "Solanum dulcamara",
    spp == "solidago rugosa" ~ "Solidago rugosa",
    spp == "solidago sempervirens" ~ "Solidago sempervirens",
    spp == "spiraea alba" ~ "Spiraea alba",
    spp == "stellaria media" ~ "Stellaria media",
    spp == "Strawberry" ~ "Fragaria virginiana",
    spp == "symphotrichum novae-angliae" ~ "Symphotrichum novae-angliae",
    spp == "taraxacum officinale" ~ "Taraxacum officinale",
    spp == "taraxacum officionale" ~ "Taraxacum officinale",
    spp == "trifolium repens" ~ "Trifolium repens",
    spp == "urtica dioica" ~ "Urtica dioica",
    spp == "vaccinium angustifolium" ~ "Vaccinium angustifolium",
    spp == "valeriana officionalis" ~ "Valeriana officinalis",
    spp == "Valeriana officionalis" ~ "Valeriana officinalis",
    spp == "valeriana officinalis" ~ "Valeriana officinalis",
    spp == "veronica arvensis" ~ "Veronica arvensis",
    TRUE ~ spp
  ))
print(island_veg_all)
```
-
```{r insert IDs (veg)}
island_veg_all <- island_veg_all %>% 
  mutate(spp = case_when(
    spp == "Min lat" ~ "Moehringia lateriflora",
    spp == "Calamagrostis #8" ~ "Poa pratensis",
    spp == "grass #54" ~ "Poa pratensis",
    spp == "Poa pratensis (#16)" ~ "Poa pratensis",
    spp == "Poa pratensis #16)" ~ "Poa pratensis",
    spp == "Poa spp #19" ~ "Alopecurus pratensis",
    spp == "elymus #33" ~ "Elymus trachycaulus",
    spp == "grass #74" ~ "Agrostis stolonifera",
    spp == "Poa #3" ~ "Poa trivialis",
    spp == "#53" ~ "Capsella bursa-pastoris",
    spp == "Poaceae #84" ~ "Elymus repens",
    spp == "Sedge #82" ~ "Carex silicea",
    spp == "Dock sp #45" ~ "Rumex britannica",
    spp == "dock sp #51" ~ "Rumex crispa",
    spp == "Poaceae #22" ~ "Anthoxanthum odoratum",
    spp == "Vetch #99" ~ "Lathyrus palustris",
    TRUE ~ spp
  ))
```
-
```{r create-veg-types}
island_veg_all$type <- "NA"
```
-
```{r assign-veg-types}
island_veg_all <- island_veg_all %>%
  # select(location:percent_cover_range) %>%
   mutate(type = case_when(
    spp == "Achillea millefolium" ~ "herb",
    spp == "Agrostis scabra" ~ "gram",
    spp == "Agrostis stolonifera" ~ "gram",
    spp == "Arctium sp." ~ "herb",
    spp == "Calystegia sepium" ~ "vine",
    spp == "Carex nigra" ~ "gram",
    spp == "Cerastium arvense" ~ "herb",
    spp == "Cerastium fontanum" ~ "herb",
    spp == "Cerastium fontanum" ~ "herb",
    spp == "Cirsium arvense" ~ "herb",
    spp == "Cirsium vulgare" ~ "herb",
    spp == "Cornus canadensis" ~ "herb",
    spp == "Cuscuta gronovii" ~ "vine",
    spp == "Doellingeria umbellata" ~ "herb",
    spp == "Festuca rubra" ~ "gram",
    spp == "Fragaria virginiana" ~ "herb",
    spp == "Hieracium caespitosum" ~ "herb",
    spp == "Impatiens capensis" ~ "herb",
    spp == "Iris versicolor" ~ "herb",
    spp == "Lathyrus japonicus" ~ "vine",
    spp == "Lathyrus palustris" ~ "vine",
    spp == "Luzula multiflora" ~ "gram",
    spp == "Maianthemum canadense" ~ "herb",
    spp == "Matricaria discoidea" ~ "herb",
    spp == "Myrica gale" ~ "shrub",
    spp == "Osmundastrum cinnamomeum" ~ "herb",
    spp == "Plantago major" ~ "herb",
    spp == "Poa pratensis" ~ "gram",
    spp == "Poa sp." ~ "gram",
    spp == "Potentilla anserina" ~ "herb",
    spp == "Potentilla norvegica" ~ "herb",
    spp == "Potentilla simplex" ~ "herb",
    spp == "Ranunculus acris" ~ "herb",
    spp == "Rubus hispidus" ~ "vine",
    spp == "Rubus idaeus" ~ "shrub",
    spp == "Rumex acetosella" ~ "herb",
    spp == "Rumex longifolius" ~ "herb",
    spp == "Rumex sp." ~ "herb",
    spp == "Rubus shrub spp. (#32)" ~ "shrub",
    spp == "Scutellaria galericulata" ~ "herb",
    spp == "Solanum dulcamara" ~ "vine",
    spp == "Solidago rugosa" ~ "herb",
    spp == "Solidago sempervirens" ~ "herb",
    spp == "Spiraea alba" ~ "shrub",
    spp == "Stellaria media" ~ "herb",
    spp == "Symphotrichum novae-angliae" ~ "herb",
    spp == "Taraxacum officinale" ~ "herb",
    spp == "Trifolium repens" ~ "herb",
    spp == "Urtica dioica" ~ "herb",
    spp == "Vaccinium angustifolium" ~ "shrub",
    spp == "Valeriana officinalis" ~ "herb",
    spp == "Veronica arvensis" ~ "herb",
    TRUE ~ type
  ))
```
##visualize the dataset
```{r find-unique-species-veg}
unique <- unique(island_veg_all$spp)
sort(unique)
```
-
```{r visualize-missing-veg-data, fig.alt = "Colored block showing missing values in the island_veg_all data where there are missing plant ids and a few missing values for percent cover."}

visdat::vis_dat(island_veg_all)
```
-
```{r visualize-percent-missing-veg-data, fig.alt = "Gray and black block showing the percentage of missing values in each column of the island_veg_all data where 15.3 percent of all of values are missing. Most of these are the OID and notes columns, and some are the species IDs and percent cover."}

visdat::vis_miss(island_veg_all)
```
-
```{r visualize-missing-variables-veg-data, fig.alt = "Bar plot of missing values in the island_veg_all data where the OID columns have the most missing values followed by the notes and species ID."}
naniar::gg_miss_var(island_veg_all)
```
-
##create veg shape file from csv
```{r create-veg-shape-file}
# #removing columns that are entirely na
# island_veg_no_na <- island_veg_all %>%
#   select(location, date, gps, cover, height, common, spp_code, spp, notes, POINT_X, POINT_Y, height_class, height_class_range, percent_cover_range, type)
#   
# #removing gdi, because it has no coordinates
# island_veg_no_na_2 <- filter(island_veg_no_na, location != "gdi")
# 
# #creating a vegetation shape file
# island_veg_shape <- st_as_sf(island_veg_no_na_2, coords = c(10:11))

island_veg_shape <- st_as_sf(island_veg_all, coords = c(13:14))

```
-
##create veg color palettes
```{r assign-colors-to-percent-cover}
cover_color <- c("0-0.01" = "#dadaeb", 
                 "0.06-0.25" = "#bcbddc", 
                 "0.26-0.50" = "#9e9ac8", 
                 "0.51-0.75" = "#756bb1", 
                 "0.76-1.00" = "#54278f") 
```
-
```{r assign-colors-to-percent-cover-for-leaflet}
cover_color_leaf <- colorFactor(palette = c("#dadaeb", "#bcbddc", "#9e9ac8", "#756bb1", "#54278f"), 
                           domain = c("0-0.01", "0.06-0.25", "0.26-0.50", "0.51-0.75", "0.76-1.00"),
                           ordered = TRUE)
```
-
```{r assign-colors-to-height-classes}
height_color <- c("0-0.15" = "#ffffb2", 
                  "0.15-0.30" = "#fecc5c", 
                  "0.30-1" = "#fd8d3c", 
                  "1-1.5" = "#f03b20", 
                  "1.5+" = "#bd0026")
```
-
```{r assign-colors-to-height-classes-for-leaflet}
height_color_leaf <- colorFactor(palette = c("#ffffb2", "#fecc5c", "#fd8d3c", "#f03b20", "#bd0026"), 
                            domain = c("0-0.15", "0.15-0.30", "0.30-1", "1-1.5", "1.5+"),
                            ordered = TRUE)
```
-

###VISUALIZING

#Visualize bird data
```{r bird-species-bar-plot-by-island, fig.alt = "Bar plots of seabird species on Great Duck Island, Heron, Schoodic, and Shabby where herring gulls are the primary species on all four islands. Heron only has herring gulls, while the other islands also have common ediers and black-backed gulls."}
 
island_bird %>%
  ggplot(mapping = aes(y = species),
  na.rm = TRUE) +
  geom_bar(fill = "#21918c") +
  facet_wrap(~location, nrow = 1) +
  theme(axis.text.y = element_text(vjust = 10)) +
  theme_minimal() + 
  labs(title = "Seabird disribution by island",
       x = "Number of occurences",
       y = "Bird species")

#ggsave("bird_species_by_island_number.jpg")
```
-
```{r species-percentage-by-island, fig.alt = "Bar plot of the proportion of each species on each island where herring gulls are the largest percentage of total birds on Great Duck Island, Heron, Schoodic, and Shabby. Shabby has the lostest proportion of herring gulls with just under 70 percent, and heron has the highest percentage at 100 percent."}

island_bird %>%
  ggplot(mapping = aes(x = location, fill = (species))) + 
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Bird species on each island",
       x = "Island",
       y = "Proportion of species",
       fill = "Bird species") +
  theme_minimal()

#ggsave("bird_species_proportion.jpg")
```
-
##geom_sf maps of species on each island
```{r ggplot-map-of-nests-on-gdi, fig.width=20, dev='jpeg', fig.alt = "Latitude and longitude map of seabird nests shown by species on Great Duck Island where most birds nest in two colonies on the northern and southern shores of the island."}

island_bird %>%
filter(location == "gdi") %>% #filter for one island for better vis
ggplot() +
  geom_sf(aes(color = species)) + #spacial geom
  scale_color_manual(values = colors) + #use same color palette so species match on each plot
  labs(title = "Seabird nests on Great Duck Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-heron, fig.width=20, dev='jpeg', fig.alt = "Latitude and longitude map of seabird nests by species on Heron Island where herring gulls are concentrated on the south shore and spaced out to the north."}

island_bird %>%
filter(location == "heron") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-schoodic, fig.width=20, dev='jpeg', fig.alt = "Latitude and longitude map of seabird nests by species on Schoodic Island where most seabirds are nesting on the northern side of the Island, including some black-backed gulls and eiders."}

island_bird %>%
filter(location == "schoodic") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-
```{r ggplot-map-of-nests-on-shabby, fig.width=20, dev='jpeg', fig.alt = "Latitude and longitude map of seabird nests by species on Shabby Island where nests are spread accross the Island with more eiders and black-backed gulls towards the center and herring gulls on the shores."}

island_bird %>%
filter(location == "shabby") %>%
ggplot() +
  geom_sf(aes(color = species)) +
  scale_color_manual(values = colors)
  labs(title = "Seabird nests on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Species")
```
-

#Visualize vegetation data
```{r species-barplot-by-island, fig.height=10, dev='jpeg', fig.alt = "Bar plots of species quantities on Great Duck, Heron, Schoodic, and Shabby Islands where there is no obvious dominant species between the islands."}

island_veg_all %>%
  filter(spp != "unknown") %>% #filtering out all of the unknown species
  ggplot(mapping = aes(y = spp)
    ) +
  geom_bar(fill = "#21918c") +
  facet_wrap(~location, nrow = 1) +
  theme(axis.text.y = element_text(vjust = 10)) +
  theme_minimal() + 
  labs(title = "Vegetation frequency by island",
       x = "Number of occurences",
       y = "Scientific names")

#ggsave("plant_species_by_island_number.jpg")

```
-
```{r height-class-distribution-by-island-bar-plot, fig.alt = "Bar plot showing the proportion of each vegetation height class on each of the four study islands where Heron and Shabby have the shortest overall vegetation structure and Great Duck is the only island with a plot in the 1.5 meters and over height class."}

island_veg_all %>%
  ggplot(mapping = aes(x = location, fill = fct_rev(height_class_range))) + #reversing height class so the tallest appears on top
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Height classes on each island",
       x = "Island",
       y = "Proportion of height class",
       fill = "Plant height classes (m)") +
  theme_minimal()

#ggsave("plant_height_proportion.jpg")
```
-
```{r hpercent-cover-distribution-by-island-bar-plot, fig.alt = "Bar plot of percent vegetation cover on each plot for all four study islands where most islands are 75 percent or more in the highest percent cover range and Shabby is only about 50 percent in the highest range."}

island_veg_all %>%
  ggplot(mapping = aes(x = location, fill = fct_rev(percent_cover_range))) + 
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Percent cover on each island",
       x = "Island",
       y = "Proportion of percent cover",
       fill = "Percent plant cover on plot") +
  theme_minimal()

#ggsave("plant_cover_proportion.jpg")
```
-
```{r veg-types-by-island, fig.alt = "Bar plot of proportion of vegetation types on each of the four study islands where Heron, Schoodic, and Shabby have a high proportion of NAs and Great Duck and Schoodic are primarily graminoids (grass like plants)."}

island_veg_all %>% 
  ggplot(mapping = aes(x = location, fill = type)) +
  geom_bar(position = "fill") +
  scale_fill_viridis_d() +
  labs(title = "Veg types on each island",
       x = "Island",
       y = "Proportion on height class",
       fill = "Plant types") +
  theme_minimal()

#ggsave("plant_type_proportion.jpg")
```
-
```{r species richness representation, fig.alt = "Ridge plot of the density species richness per plot where Great Duck and Heron have a biomodal distribution at 0 and 7.5-10, plots on Shabby had an average of 2.5 species per plot, and Schoodic have an average of 2.75 species per plot."}

#ridge plot of # of species in each plot by island
iva <- island_veg_all %>%
  group_by(gps) %>%
  mutate(spp = replace_na(spp, "1")) %>%
   mutate(numSp = n_distinct(spp)) %>%
   mutate(NSp = if_else(spp != 1, numSp, 0)) %>%
  select(-numSp) %>%
  aggregate(NSp~gps+location, FUN = "mean")

ggplot(iva, aes(NSp, location, fill = fct_rev(location))) +
  geom_density_ridges(alpha = 0.5, bandwidth = .6) +
  xlim(0,15) +
  scale_fill_viridis_d() +
  labs(title = "Species Richness per Plot, by Island",
       y = "Island",
       x = "# of Species/Plot",
       fill = "Island"
  ) +
  theme_minimal()

#ggsave("richness_ridge_plot.jpg")
```
-
##geom_point maps of the percent cover on each island
```{r ggplot-geom_point-map-percent-cover-on-heron, fig.alt = "Latitude and longitude map of vegetation plots by percent cover on Heron Island where the lower percent cover plots appear to be closer the the shores."}

island_veg_all %>%
  filter(location == "heron") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) + #geom point because data doesn't read as geometry. Rename column?
  scale_color_manual(values = cover_color) + #color scale for percent cover, darker is more cover
  scale_y_continuous(labels = ~ paste0(.x, "°N")) + #add lat/long symbols to axis
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
-
```{r ggplot-geom_point-map-percent-cover-on-schoodic, fig.alt = "Latitude and longitude map of vegetation maps by percent cover on Schoodic Island where most of the plots have a high percent cover."}

island_veg_all %>%
  filter(location == "schoodic") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) +
  scale_color_manual(values = cover_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
-
```{r ggplot-geom_point-map-percent-cover-on-shabby, fig.alt = "Latitude and longitude map of vegetation plots by percent cover on Shabby Island where the percent cover makes no obvious pattern."}

island_veg_all %>%
  filter(location == "shabby") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = percent_cover_range)) +
  scale_color_manual(values = cover_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Percent cover on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Percent cover")
```
-
##geom_point maps of the height classes on each island
```{r ggplot-geom_point-map-height-on-heron, fig.alt = "Latitude and longitude map of vegetation plots by height class on Heron Island where there is no obvious pattern to the height classes."}

island_veg_all %>%
  filter(location == "heron") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) + #rev height range so it goes low to high
  scale_color_manual(values = height_color) + #special color scale for height
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Heron Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-
```{r ggplot-geom_point-map-height-on-schoodic, fig.alt = "Latitude and longitude map of vegetation plots by height class on Schoodic Island where there is no obvious pattern to the heigh classes."}

island_veg_all %>%
  filter(location == "schoodic") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) +
  scale_color_manual(values = height_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Schoodic Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-
```{r ggplot-geom_point-map-height-on-shabby, fig.alt = "Latitude and longitude map of vegetation plots by height class on Shabby Island where the height classes make no obvious pattern."}

island_veg_all %>%
  filter(location == "shabby") %>%
  ggplot() +
  geom_point(aes(POINT_X, POINT_Y, color = fct_rev(height_class_range))) +
  scale_color_manual(values = height_color) +
  scale_y_continuous(labels = ~ paste0(.x, "°N")) +
  scale_x_continuous(labels = ~ paste0(.x, "°W")) +
  labs(title = "Height class on Shabby Island",
       x = "Longitude",
       y = "Latitude",
       color = "Height class")
```
-
#Compare birds and vegetation

```{r change_coordinate_system_for_bird}
#update bird data to the coordinate system used in leaflet
island_bird <- st_transform(island_bird, "+init=epsg:4326")
```
-
```{r leaflet-map-of-height-and-bird-species, fig.alt = "Interactive Leaflet map of seabird nests, by species, and vegetation plots, by height class, where nesting sites may be correlated with vegetation height."}

leaflet() %>% 
  #basemap
  addProviderTiles("CartoDB.Positron") %>% 
  #setView to show all islands
  setView(lng = -68.25,
          lat = 44.25,
          zoom = 10) %>% 
  #add markers for the birds and color by species using bird_color palette
  addCircleMarkers(data = island_bird,
                   color = ~bird_color(island_bird$species),
                   weight = 1,
                   fillColor = ~bird_color(island_bird$species),
                   radius = 3,
                   opacity = .5,
                   fillOpacity = 1) %>% 
  #add markers for veg colored by height class using the height_color_leaf palette
  addCircleMarkers(data = island_veg_shape, 
                   color = "black", 
                   weight = 1, #~height_color_leaf(island_veg_shape$height_class_range), 
                   fillColor = ~height_color_leaf(island_veg_shape$height_class_range),
                   radius = 3,
                   opacity = 1,
                   fillOpacity = 1)
```
#I would like to have a representation of bird densities and then veg plots colored by height
#also veg plots represented by dominant vegetation type in colors
```{r species richness representations}
# #ridge plot of # of species in each plot by island
# iva <- island_veg_all %>%
#   group_by(gps) %>%
#   mutate(spp = replace_na(spp, "1")) %>%
#    mutate(numSp = n_distinct(spp)) %>%
#    mutate(NSp = if_else(spp != 1, numSp, 0)) %>%
#   select(-numSp) %>%
#   aggregate(NSp~gps+location, FUN = "mean")
# =======

    
```

Next steps: - veg plots represented by dominant vegetation type in colors
            - presence/absence of armature on veg
            - GIS work --> classification wizard
>>>>>>> dc9585e7f9ddfe982e9684fa1e79a99c65bb6eaa

-
```{r working code leaflet - set shapes (do not use: ugly)}

# add icon label column
island_bird_1 <- island_bird %>%
  mutate(icon = case_when(
    species == "great black-backed gull" ~ "square",
    species == "common eider" ~ "circle",
    species == "herring gull" ~ "triangle"))

# # Make a list of icons. We'll index into it based on name.
# /!\ order of icons MUST BE THE SAME as the order of the factor "group_color"
bird_icons <- iconList(
  circle <- makeIcon(iconUrl = "https://creazilla-store.fra1.digitaloceanspaces.com/emojis/45166/black-circle-emoji-clipart-md.png",
                          iconWidth = 6, iconHeight = 6),
  square <- makeIcon(iconUrl = "https://creazilla-store.fra1.digitaloceanspaces.com/emojis/45970/black-medium-small-square-emoji-clipart-md.png",
                           iconWidth = 7, iconHeight = 7),
  triangle <- makeIcon(iconUrl = "https://www.freeiconspng.com/uploads/triangle-png-28.png",
                         iconWidth = 6, iconHeight = 6)
  )
leaflet() %>%
  #basemap
  addProviderTiles("CartoDB.Positron") %>%
  #setView to show all islands
  setView(lng = -68.25,
          lat = 44.25,
          zoom = 10) %>%
      #add markers for veg colored by height class using the height_color_leaf palette
  addCircleMarkers(data = island_veg_shape, 
                   color = ~height_color_leaf(island_veg_shape$height_class_range), 
                   fillColor = ~height_color_leaf(island_veg_shape$height_class_range),
                   radius = 2,
                   opacity = 1,
                   fillOpacity = 1)   %>%
  #add markers for the birds and color by species using bird_color palette
  addMarkers(data = island_bird_1,
             icon = ~ bird_icons
                   ) 

  
  
```

